Constraining the Merger History of Primordial-Black-Hole Binaries from GWTC-3

### 2023-04-02 latex/paper.tex
update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{}
%\revised{}
%\accepted{}
%\published{}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
%\submitjournal{ApJ}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\title{Constrain the Merger History of Primordial-Black-Hole Binaries from GWTC-3}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \author[0000-0002-3309-415X]{Zhi-Qiang~You}
\author{Zhi-Qiang~You}
\email{zhiqiang.you@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

@@

The successful detection of gravitational waves (GWs) from compact binary coalescences \cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl,LIGOScientific:2021djp} has led us into a new era of GW astronomy.
According to the recently released third GW Transient Catalog (GWTC-3) \cite{LIGOScientific:2021djp} by LIGO-Virgo-KAGRA (LVK) Collaboration, there are $90$ GW events detected during the first three observing runs. Most of these events are categorized as binary black hole (BBH) mergers, and the BBHs detected by LVK have a broad mass distribution. The heaviest event, GW190521 \cite{LIGOScientific:2020iuh}, has component masses $m_1 = 85^{+21}_{-14}\Msun$ and $m_2 = 66^{+17}_{-18}\Msun$. Both masses lie within upper black hole mass gap originated from pulsation pair-instability supernovae \cite{Marchant_2019}, and current modelling places the lower cutoff of the mass gap at $\sim 50\pm 4\Msun$ \cite{Belczynski:2016jno,Marchant_2019,Farmer_2019,Farmer:2020xne,Marchant:2020haw}.
Even accounting for the uncertainties in modelling, it still implies at least $m_1$ is well within the mass gap and cannot originate directly from a stellar progenitor \cite{Anagnostou:2020umw}.
Therefore, the heavy event GW190521 greatly challenges the stellar evolution scenario of astrophysical black holes.

Besides the astrophysical black holes, another possible explanation for the LVK BBHs is the primordial black holes (PBHs) \cite{Bird:2016dcv,Sasaki:2016jop,Chen:2018czv,Liu:2018ess,Chen:2021nxo}. 
PBHs are black holes formed in the very early Universe through the gravitational collapse of the primordial density fluctuations \cite{Hawking:1971ei,Carr:1974nx}. Recently, PBHs have attracted considerable attention \cite{Garcia-Bellido:2017mdw,Carr:2017jsz,Germani:2017bcs,Liu:2019rnx,Cai:2019elf,Cai:2019bmk,DeLuca:2020sae,Vaskonen:2020lbd,DeLuca:2020agl,Hutsi:2020sol,Sasaki:2018dmp,Carr:2020gox,Carr:2020xqk,Liu:2021jnw,Wang:2022nml,Franciolini:2022tfm} because they can be not only the sources of LVK detections~\cite{Bird:2016dcv,Sasaki:2016jop}, but also candidates of cold dark matter (CDM) \cite{Carr:2016drx} and the seeds for galaxy formation~\cite{Bean:2002kx,Kawasaki:2012kn}. The formation of PBHs would inevitably accompany the production of scalar-induced GWs \cite{Saito:2008jc,Yuan:2019udt,Yuan:2019wwo,Yuan:2019fwv,Chen:2019xse,DeLuca:2019ufz,Bartolo:2018rku,Bartolo:2018evs}.
Recent studies \cite{Chen:2021nxo,Chen:2022fda} show that the BBHs from GWTC-3 are consistent with the PBH scenario, and the abundance of PBH in CDM, $\fpbh$, should be in the order of $\mathcal{O}(10^{-3})$ to explain LVK BBHs. In particular, the merger rate for GW190521 derived from the PBH model is consistent with that inferred by LVK, indicating that GW190521 can be a PBH binary \cite{DeLuca:2020sae,Chen:2021nxo}.

Accurately estimating the merger rate distribution of PBH binaries can be crucial to extract the PBH population parameters from GW data. Ref.~\cite{Liu:2019rnx} studies the multi-merger processes of PBH binaries and show that the merger history of PBH binaries may shift the mass distribution from light mass to heavy mass depending on the values of population parameters. Ref.~\cite{Wu:2020drm} then infers the population parameters of PBH binaries by accounting for the merger history effect using $10$ BBHs from GWTC-1, finding that the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries. 
In this work, we use the LVK recent released GWTC-3 data to constrain the effect of merger history on the merger rate of PBH binaries assuming all LVK BBHs are of primordial origin.
We extend the analyses of Ref.~\cite{Wu:2020drm} in several aspects. Firstly, we use GWTC-3, which expands GWTC-1 with almost six times more BBH events. The GWTC-3 events expand the mass and redshift coverage and can alleviate the statistical bias by including significantly more BBHs. Secondly, Ref.~\cite{Wu:2020drm} only considers the PBH mass functions with the log-normal and power-law forms. We do more comprehensive analyses by including the broken power-law and critical collapse PBH mass functions that were not considered in Ref.~\cite{Wu:2020drm}. It is claimed by Ref.~\cite{Deng:2021ezy} that a broken power-law can fit the GW data better than the log-normal form. Lastly, we consider the redshift distribution of the merger rate that is ignored in Ref.~\cite{Wu:2020drm}.

@@

We can now estimate the merger rate densities of PBH binaries by considering the merger history effect.
We assume that PBHs are randomly distributed in the early Universe when they decouple from the cosmic background evolution \cite{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. The two nearest PBHs would attract each other because of the gravitational interactions. These two PBHs would obtain the angular momentum from the torque of other PBHs and form a PBH binary. The binary would emit gravitational radiations and eventually merge.

We do not intend to give a detailed derivation but quote the results from Ref.~\cite{Liu:2019rnx} here.
The merger rate density from first-merger process, $\cR_1(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \int \hat{\cR}_1 \rd m_l,
\q 
where
\e 
\begin{split}
	&\hat{\cR}_1(t, m_i, m_j, m_l)
	\equiv 1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
	&\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
Similarly, the merger rate density from second-merger process, $\cR_2(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\begin{split}
	\cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
	&+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where
\e 
\begin{split}
	&\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
	=1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
	&\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
	&\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
We only consider the effect of merger history up to the second-merger process.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, of PBH binaries at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
\cR(t, m_i, m_j) =  \sum_{n=1,2} \cR_n(t, m_i, m_j),
\q 
and the total merger rate is
\e\label{Rt}
R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1,2} R_n(t),
\q 
where
\e 
R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
All the above-mentioned merger rate (density) is measured at the source frame.
We should emphasize that although $R_2(t)$ should be smaller than $R_1(t)$ as expected, $\cR_2(t, m_i, m_j)$ is not necessarily be smaller than $\cR_1(t, m_i, m_j)$ \cite{Liu:2019rnx}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbp!]
	\centering
	\begin{tabular}{lll}
		\hline\hline
		\textbf{Parameter\quad} & \textbf{Description} & \textbf{Prior} \\
		\hline
		$\fpbh$ & Abundance of PBH in CDM & log-$\mU(-4, 0)$\\
		\hline
		\multicolumn{3}{c}{Lognormal PBH mass function} \\[1pt]
		$\Mc$ & Central mass in $\Msun$. & $\mU(5, 50)$\\
		$\sigma$ & Mass width. & $\mU(0.1, 2)$\\
		\hline
		\multicolumn{3}{c}{Power-law PBH mass function} \\[1pt]
		$\Mmin$ & Lower mass cut-off in $\Msun$. & $\mU(3, 10)$\\
		$\al$ & Power-law index. & $\mU(1.05, 4)$\\
		\hline
		\multicolumn{3}{c}{Broken Power-law PBH mass function} \\[1pt]
		$M_*$ & Peak mass in $\Msun$. & $\mU(3, 15)$\\
		$\al_1$ & First power-law index. & $\mU(0, 15)$\\
		$\al_2$ & Second power-law index. & $\mU(1, 3)$\\
		\hline
		\multicolumn{3}{c}{Critical collapse (CC) PBH mass function} \\[1pt]
		$\Mf$ & Horizon mass scale in $\Msun$. & $\mU(1, 50)$\\
		$\al$ & Universal exponent. & $\mU(0, 5)$\\
		\hline
	\end{tabular}	
	\caption{\label{table:priors}Parameters and their prior distributions used in the Bayesian parameter estimations. Here, $\mU$ and log-$\mU$ denote uniform and log-uniform distributions, respectively.}
\end{table}

Given the data, $\textbf{d} = \{d_1, d_2, \cdots, d_{N_{\mathrm{obs}}}\}$, of $N_{\mathrm{obs}}$ BBH merger events, we model the total number of events as an inhomogeneous Poisson process, yielding the likelihood \cite{Loredo:2004nn,Thrane:2018qnx,Mandel:2018mve}
\begin{equation}\label{L1}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{\int \mathcal{L} (d_{i}| \theta)\, \cR_{\mathrm{pop}}(\theta|\Ld) d \theta}{\xi(\Ld)},
\end{equation}
where $N_{\exp}\equiv N_{\exp}(\Ld)$ is the expected number of detections over the timespan of observation. Here $\mathcal{L} (d_{i}|\theta)$ is the individual event likelihood for the $i$th GW event that can be derived from the individual event's posterior by reweighing with the prior on $\theta$. Here, $\xi(\Ld)$ quantifies selection biases for a population with parameters $\Ld$ and is defined by
\e\label{xi}
\xi(\Ld) = \int P_{\mathrm{det}}(\theta)\, \cR_{\mathrm{pop}}(\theta|\Ld)\, \mathrm{d} \theta,
\q
where $P_{\text{det}}(\theta)$ is the detection probability that depends on the source parameters $\theta$.
In practice, we use the simulated injections \cite{ligo_scientific_collaboration_and_virgo_2021_5546676} to estimate $\xi(\Ld)$, and \Eq{xi} can be approximated by a Monte Carlo integral over found injections \cite{LIGOScientific:2021psn}
\begin{equation}
	\xi(\Ld) \approx \frac{1}{N_{\mathrm{inj}}} \sum_{j=1}^{N_{\text {found }}} \frac{\cR_{\mathrm{pop}}(\theta_{j} | \Ld)}{p_{\mathrm{draw}}(\theta_j)},
\end{equation}
where $N_{\text{inj}}$ is the total number of injections, $N_{\text{found}}$ is the number of successfully detected injections, and $p_{\mathrm{draw}}$ is the probability density function from which the injections are drawn.
Using the posterior samples from each event, we estimate the hyper-likelihood \eqref{L1} as
\begin{equation}\label{L2}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{1}{\xi(\Ld)} \left\langle \frac{\cR_{\mathrm{pop}}(\theta|\Ld)}{d_L^2(z)} \right\rangle,
\end{equation}
where $\langle\cdots\rangle$ denotes the weighted average over posterior samples of $\theta$. The denominator $d_L^2(z)$ is the standard priors used in the LVK analysis of individual events where $d_L$ is the luminosity distance.

In this work, we incorporate the PBH population distribution \eqref{cR} into the \texttt{ICAROGW} \cite{Mastrogiovanni:2021wsd} package to estimate the likelihood function \eqref{L2}, and use \texttt{dynesty} \cite{Speagle:2019ivv} sampler called from \texttt{Bilby} \cite{Ashton:2018jfp,Romero-Shaw:2020owr} to sample over the parameter space. We use the GW events from GWTC-3 by discarding events with false alarm rate larger than 1 yr$^{-1}$ and events with the secondary component mass smaller than $3\Msun$ to avoid contamination from putative events involving neutron stars following Ref.~\cite{DeLuca:2021wjr}. A total of $69$ GW events from GWTC-3 meet this criteria and the posterior samples of these BBHs are publicly available from Ref.~\cite{ligo_scientific_collaboration_and_virgo_2021_5655785}.

@@

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=\linewidth]{post-log.pdf}
	\caption{\label{posterior-log}The marginalized one- and two-dimensional posterior distributions for hyper-parameters $\{M_c, \s, \fpbh\}$ in the log-normal mass function inferred from GWTC-3. The blue color denotes the results from the first merger only, while the orange denotes the results from both the first and second mergers. The contours represent the $1\sigma$, $2\sigma$ and $3\sigma$ credible regions, respectively.}
\end{figure}

@@

Using $69$ BBHs from GWTC-3 and performing the hierarchical Bayesian inference, we obtain $M_c = 17.3^{+2.2}_{-2.0} \Msun$, $\s = 0.71^{+0.10}_{-0.08}$, and $\fpbh = 1.8^{+0.3}_{-0.3} \times 10^{-3}$. In this work, we present results with median value and 90\% equal-tailed credible intervals. The posteriors for the hyper-parameters $\Ld = \{M_c, \s, \fpbh\}$ are shown in \Fig{posterior-log}. Note that we get a larger value of $M_c$ than that inferred from GWTC-1 in Ref.~\cite{Wu:2020drm} because GWTC-3 contains heavier BHs than those from GWTC-1. From \Eq{Rt}, we also infer the local merger rate as $R(t_0) = 41^{+16}_{-12} \gpcyr$. The results of local merger rate and abundance of PBHs are consistent with the previous estimations \cite{Sasaki:2016jop,Ali-Haimoud:2017rtz,Chen:2018czv,Chen:2018rzo,Chen:2019irf,Wu:2020drm,Chen:2021nxo,Chen:2022fda}, confirming that CDM cannot be dominated by the stellar-mass PBHs.

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=\linewidth]{ratio-log.pdf}
	\caption{\label{ratio-log}The ratio of merger rate density from second merger to that from first merger, $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$, as a function of component masses for the log-normal mass function. We have fixed the hyper-parameters $\{M_c, \s, \fpbh\}$ to their best-fit values.}
\end{figure}

In \Fig{ratio-log}, we show the ratio of merger rate density from second merger to the one from first merger, namely $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$, by fixing the hyper-parameters $\{M_c, \s, \fpbh\}$ to their best-fit values. It can be seen that the second merger provides more contribution to the total merger rate density as component mass increases. Even though $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$ can reach as high as $\gtrsim 10\%$, the ratio of merger rate from second merger to that from first merger is $R_2(t_0)/R_1(t_0) = 1.0^{+0.2}_{-0.1}\%$ and is negligible. This is because the major contribution to the merger rate is from the masses less than $50\Msun$, and the correction is negligible in this mass range. Therefore the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries.

>>>>

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\title{Constraining the Merger History of Primordial-Black-Hole Binaries from GWTC-3}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \author[0000-0002-3309-415X]{Zhi-Qiang~You}
\author{Zhi-Qiang~You}
\email{Corresponding author: zhiqiang.you@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

@@

The successful detection of gravitational waves (GWs) from compact binary coalescences \cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl,LIGOScientific:2021djp} has led us into a new era of GW astronomy.
According to the recently released third GW Transient Catalog (GWTC-3) \cite{LIGOScientific:2021djp} by LIGO-Virgo-KAGRA (LVK) Collaboration, there are $90$ GW events detected during the first three observing runs. Most of these events are categorized as binary black hole (BBH) mergers, and the BBHs detected by LVK have a broad mass distribution. The heaviest event, GW190521 \cite{LIGOScientific:2020iuh}, has component masses $m_1 = 85^{+21}_{-14}\Msun$ and $m_2 = 66^{+17}_{-18}\Msun$. Both masses lie within upper black hole mass gap originated from pulsation pair-instability supernovae \cite{Marchant_2019}, and current modelling places the lower cutoff of the mass gap at $\sim 50\pm 4\Msun$ \cite{Belczynski:2016jno,Marchant_2019,Farmer_2019,Farmer:2020xne,Marchant:2020haw}.
Even accounting for the statistical uncertainties, it still implies at least $m_1$ is well within the mass gap and cannot originate directly from a stellar progenitor \cite{Anagnostou:2020tta}.
Therefore, the heavy event GW190521 greatly challenges the stellar evolution scenario of astrophysical black holes.

Besides the astrophysical black holes, another possible explanation for the LVK BBHs is the primordial black holes (PBHs) \cite{Bird:2016dcv,Sasaki:2016jop,Chen:2018czv,Liu:2018ess,Chen:2021nxo}. 
PBHs are black holes formed in the very early Universe through the gravitational collapse of the primordial density fluctuations \cite{Hawking:1971ei,Carr:1974nx}. Recently, PBHs have attracted considerable attention \cite{Garcia-Bellido:2017mdw,Carr:2017jsz,Germani:2017bcs,Liu:2019rnx,Cai:2019elf,Cai:2019bmk,DeLuca:2020sae,Vaskonen:2020lbd,DeLuca:2020agl,Hutsi:2020sol,Sasaki:2018dmp,Carr:2020gox,Carr:2020xqk,Liu:2021jnw,Wang:2022nml,Franciolini:2022tfm} because they can be not only the sources of LVK detections~\cite{Bird:2016dcv,Sasaki:2016jop}, but also candidates of cold dark matter (CDM) \cite{Carr:2016drx} and the seeds for galaxy formation~\cite{Bean:2002kx,Kawasaki:2012kn}. The formation of PBHs would inevitably accompany the production of scalar-induced GWs \cite{Saito:2008jc,Cai:2018dig,Yuan:2019udt,Yuan:2019wwo,Yuan:2019fwv,Chen:2019xse,DeLuca:2019ufz,Bartolo:2018rku,Bartolo:2018evs}.
Recent studies \cite{Chen:2021nxo,Chen:2022fda} show that the BBHs from GWTC-3 are consistent with the PBH scenario, and the abundance of PBH in CDM, $\fpbh$, should be in the order of $\mathcal{O}(10^{-3})$ to explain LVK BBHs. In particular, the merger rate for GW190521 derived from the PBH model is consistent with that inferred by LVK, indicating that GW190521 can be a PBH binary \cite{DeLuca:2020sae,Chen:2021nxo}.

Accurately estimating the merger rate distribution of PBH binaries can be crucial to extract the PBH population parameters from GW data. Ref.~\cite{Liu:2019rnx} studies the multi-merger processes of PBH binaries and show that the merger history of PBH binaries may shift the mass distribution from light mass to heavy mass depending on the values of population parameters. Ref.~\cite{Wu:2020drm} then infers the population parameters of PBH binaries by accounting for the merger history effect using $10$ BBHs from GWTC-1, finding that the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries. 
In this work, we use the LVK recent released GWTC-3 data to constrain the effect of merger history on the merger rate of PBH binaries assuming all LVK BBHs are of primordial origin.
We extend the analyses of Ref.~\cite{Wu:2020drm} in several aspects. Firstly, we use a purified subset of GWTC-3, which expands GWTC-1 with almost six times more BBH events. The GWTC-3 events expand the mass and redshift coverage and can alleviate the statistical bias by including significantly more BBHs. Secondly, Ref.~\cite{Wu:2020drm} only considers the PBH mass functions with the log-normal and power-law forms. We do more comprehensive analyses by including the broken power-law and critical collapse PBH mass functions that were not considered in Ref.~\cite{Wu:2020drm}. It is claimed by Ref.~\cite{Deng:2021ezy} that a broken power-law can fit the GW data better than the log-normal form. Lastly, we consider the redshift distribution of the merger rate that is ignored in Ref.~\cite{Wu:2020drm}.
The aforementioned reasons have inspired us to explore the possibility that the heavy black holes detected by LVK have been formed, at least in part, through second-generation mergers. This is because the second-merger process has the potential to increase the mass distribution to a higher value. A precise assessment of the influence of second-generation mergers on mass distribution demands a meticulous analysis of the data, as has been conducted in this study.

@@

We can now estimate the merger rate densities of PBH binaries by considering the merger history effect.
We assume that PBHs are randomly distributed following a spatial Poisson distribution in the early Universe when they decouple from the cosmic background evolution \cite{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. The two nearest PBHs would attract each other because of the gravitational interactions. These two PBHs would obtain the angular momentum from the torque of other PBHs and form a PBH binary after decoupling from the cosmic expansion. The binary would emit gravitational radiations and eventually merge.

We do not intend to give a detailed derivation but quote the results from Ref.~\cite{Liu:2019rnx} here.
The merger rate density from first-merger process, $\cR_1(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \int \hat{\cR}_1 \rd m_l,
\q 
where $m_i$ and $m_j$ are the masses of the merging binary, $m_l$ is the mass of the third black hole that is closest to the merging binary, and
\e 
\begin{split}
	&\hat{\cR}_1(t, m_i, m_j, m_l)
	\equiv 1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
	&\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
Here, $t$ is the cosmic time, and $t_0$ is the present cosmic time.
Similarly, the merger rate density from second-merger process, $\cR_2(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\begin{split}
	\cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
	&+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where $m_e$ is the mass of the fourth black hole that is closest to the merging binary, and
\e 
\begin{split}
	&\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
	=1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
	&\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
	&\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
We only consider the effect of merger history up to the second-merger process. We have verified that the fraction of the third merger rate over the second merger rate is less than $0.005$.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, of PBH binaries at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
\cR(t, m_i, m_j) =  \sum_{n=1,2} \cR_n(t, m_i, m_j),
\q 
and the total merger rate is
\e\label{Rt}
R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1,2} R_n(t),
\q 
where
\e 
R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
All the above-mentioned merger rate (density) is measured at the source frame.
We should emphasize that although $R_2(t)$ should be smaller than $R_1(t)$ as expected, $\cR_2(t, m_i, m_j)$ is not necessarily be smaller than $\cR_1(t, m_i, m_j)$ \cite{Liu:2019rnx}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbp!]
	\centering
	\begin{tabular}{lll}
		\hline\hline
		\textbf{Parameter\quad} & \textbf{Description} & \textbf{Prior} \\
		\hline
		$\fpbh$ & Abundance of PBH in CDM & log-$\mU(-4, 0)$\\
		\hline
		\multicolumn{3}{c}{Lognormal PBH mass function} \\[1pt]
		$\Mc$ & Central mass in $\Msun$. & $\mU(5, 50)$\\
		$\sigma$ & Mass width. & $\mU(0.1, 2)$\\
		\hline
		\multicolumn{3}{c}{Power-law PBH mass function} \\[1pt]
		$\Mmin$ & Lower mass cut-off in $\Msun$. & $\mU(3, 10)$\\
		$\al$ & Power-law index. & $\mU(1.05, 4)$\\
		\hline
		\multicolumn{3}{c}{Broken Power-law PBH mass function} \\[1pt]
		$m_*$ & Peak mass in $\Msun$. & $\mU(5, 50)$\\
		$\al_1$ & First power-law index. & $\mU(0, 3)$\\
		$\al_2$ & Second power-law index. & $\mU(1, 10)$\\
		\hline
		\multicolumn{3}{c}{Critical collapse (CC) PBH mass function} \\[1pt]
		$\Mf$ & Horizon mass scale in $\Msun$. & $\mU(1, 50)$\\
		$\al$ & Universal exponent. & $\mU(0, 5)$\\
		\hline
	\end{tabular}	
	\caption{\label{table:priors}Parameters and their prior distributions used in the Bayesian parameter estimations. Here, $\mU$ and log-$\mU$ denote uniform and log-uniform distributions, respectively.}
\end{table}

Given the data, $\textbf{d} = \{d_1, d_2, \cdots, d_{N_{\mathrm{obs}}}\}$, of $N_{\mathrm{obs}}$ BBH merger events, we model the total number of events as an inhomogeneous Poisson process, yielding the likelihood \cite{Loredo:2004nn,Thrane:2018qnx,Mandel:2018mve}
\begin{equation}\label{L1}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{\int \mathcal{L} (d_{i}| \theta)\, \cR_{\mathrm{pop}}(\theta|\Ld) d \theta}{\xi(\Ld)},
\end{equation}
where $N_{\exp}\equiv N_{\exp}(\Ld)$ is the expected number of detections over the timespan of observation. Here $\mathcal{L} (d_{i}|\theta)$ is the individual event likelihood for the $i$th GW event that can be derived from the individual event's posterior by reweighing with the prior on $\theta$. Here, $\xi(\Ld)$ quantifies selection biases for a population with parameters $\Ld$ and is defined by
\e\label{xi}
\xi(\Ld) = \int P_{\mathrm{det}}(\theta)\, \cR_{\mathrm{pop}}(\theta|\Ld)\, \mathrm{d} \theta,
\q
where $P_{\text{det}}(\theta)$ is the detection probability that depends on the source parameters $\theta$.
In practice, we use the simulated injections \cite{ligo_scientific_collaboration_and_virgo_2021_5546676} to estimate $\xi(\Ld)$, and \Eq{xi} can be approximated by a Monte Carlo integral over found injections \cite{KAGRA:2021duu}
\begin{equation}
	\xi(\Ld) \approx \frac{1}{N_{\mathrm{inj}}} \sum_{j=1}^{N_{\text {found }}} \frac{\cR_{\mathrm{pop}}(\theta_{j} | \Ld)}{p_{\mathrm{draw}}(\theta_j)},
\end{equation}
where $N_{\text{inj}}$ is the total number of injections, $N_{\text{found}}$ is the number of successfully detected injections, and $p_{\mathrm{draw}}$ is the probability density function from which the injections are drawn.
Using the posterior samples from each event, we estimate the hyper-likelihood \eqref{L1} as
\begin{equation}\label{L2}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{1}{\xi(\Ld)} \left\langle \frac{\cR_{\mathrm{pop}}(\theta|\Ld)}{d_L^2(z)} \right\rangle,
\end{equation}
where $\langle\cdots\rangle$ denotes the weighted average over posterior samples of $\theta$. The denominator $d_L^2(z)$ is the standard priors used in the LVK analysis of individual events where $d_L$ is the luminosity distance.

In this work, we incorporate the PBH population distribution \eqref{cR} into the \texttt{ICAROGW} \cite{Mastrogiovanni:2021wsd} package to estimate the likelihood function \eqref{L2}, and use \texttt{dynesty} \cite{Speagle:2019ivv} sampler called from \texttt{Bilby} \cite{Ashton:2018jfp,Romero-Shaw:2020owr} to sample over the parameter space. We use the GW events from GWTC-3 by discarding events with false alarm rate larger than 1 yr$^{-1}$ and events with the secondary component mass smaller than $3\Msun$ to avoid contamination from putative events involving neutron stars following Ref.~\cite{DeLuca:2021wjr}. A total of $69$ GW events from GWTC-3 meet these criteria and the posterior samples of these BBHs are publicly available from Ref.~\cite{ligo_scientific_collaboration_and_virgo_2021_5655785}.

@@

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=\linewidth]{post-log.pdf}
	\caption{\label{posterior-log}The marginalized one- and two-dimensional posterior distributions for hyper-parameters $\{M_c, \s, \fpbh\}$ in the log-normal mass function inferred from GWTC-3. The blue color denotes the results from the first merger only, while the orange denotes the results from both the first and second mergers. The contours represent the $1\sigma$, $2\sigma$, and $3\sigma$ credible regions, respectively.}
\end{figure}

@@

Using $69$ BBHs from GWTC-3 and performing the hierarchical Bayesian inference, we obtain $M_c = 17.3^{+2.2}_{-2.0} \Msun$, $\s = 0.71^{+0.10}_{-0.08}$, and $\fpbh = 1.8^{+0.3}_{-0.3} \times 10^{-3}$. In this work, we present results with median value and 90\% equal-tailed credible intervals. The posteriors for the hyper-parameters $\Ld = \{M_c, \s, \fpbh\}$ are shown in \Fig{posterior-log}. Note that we get a larger value of $M_c$ than that inferred from GWTC-1 in Ref.~\cite{Wu:2020drm} because GWTC-3 contains heavier BHs than those from GWTC-1. From \Eq{Rt}, we also infer the local merger rate as $R(t_0) = 41^{+16}_{-12} \gpcyr$. The results of local merger rate and abundance of PBHs are consistent with the previous estimations \cite{Sasaki:2016jop,Ali-Haimoud:2017rtz,Chen:2018czv,Chen:2018rzo,Chen:2019irf,Wu:2020drm,Chen:2021nxo,Chen:2022fda,Zheng:2022wqo}, confirming that CDM cannot be dominated by the stellar-mass PBHs.

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=\linewidth]{ratio-log.pdf}
	\caption{\label{ratio-log}The ratio of merger rate density from the second merger to that from the first merger, $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$, as a function of component masses for the log-normal mass function. We have fixed the hyper-parameters $\{M_c, \s, \fpbh\}$ to their best-fit values.}
\end{figure}

In \Fig{ratio-log}, we show the ratio of merger rate density from the second merger to the one from the first merger, namely $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$, by fixing the hyper-parameters $\{M_c, \s, \fpbh\}$ to their best-fit values. It can be seen that the second merger provides more contribution to the total merger rate density as component mass increases. Even though $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$ can reach as high as $\gtrsim 10\%$, the ratio of merger rate from second merger to that from the first merger is $R_2(t_0)/R_1(t_0) = 1.0^{+0.2}_{-0.1}\%$ and is negligible. This is because the major contribution to the merger rate is from the masses less than $50\Msun$, and the correction is negligible in this mass range. Therefore the effect of merger history can be safely ignored when estimating the merger rate of PBH binaries.
